Unified visualization of seasonal influenza evolution in the laboratory and in nature.

### 2019-06-11 paper/paper.tex
Use acmart document class without ACM-style headers and footers
This removes nonsensical conference in Washington DC from headers and footers,
since this is not truly a conference proceeding.
Fix figure layout
Use `\columnwidth` directive to scale figures to fit in a column and crop
figures to remove extraneous whitespace.
Removed whitespace
3-D to 3D

%%
%% This is file `sample-sigchi.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigchi')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigchi.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigchi]{acmart}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Unified visualization of seasonal influenza evolution in the laboratory and in nature.
}

@@

Despite the value of these data, they are not intuitive to interpret.
Sequencing produces linear sequences, yet the protein conformation may have amino acid residues that are close together in 3D space that are quite distant from each other in linear sequence space.
Thus it can be hard to interpret how changes visible on a sequence map onto the protein structure.
Additionally, as DMS experiments represent controlled laboratory experiments looking at infection in cell culture, there are questions regarding how well DMS results reflect the patterns of influenza evolution that we observe in nature.
Here, we introduce an interactive visualization platform that allows unified display of DMS differential selection data, 3-D protein structure, and information about amino acid preferences observed in nature, as unified by their common variable: site in the the sequence.

@@

@@

This visualization platform links three data views together by their common shared element, site in the protein.
The three data streams are: 1) DMS differential selection value, which is a quantitative measure of the preference for a certain amino acid at a specific site in the protein, 2) the natural amino acid frequencies of a site, a quantitative variable the describes what proportion of all sampled strains of influenza at time T have a specific amino acid at a site, and 3) the 3-D structure of protein itself, which takes linear sequence space and folds that up to reflect where certain sites of the sequence fall on the protein structure.
We describe design decisions about how to visualize these data streams below.
Our visualization was implemented using D3 \cite{bostock2011d3}

\begin{figure}[H]
	\includegraphics[width=1.0\textwidth]{viz-overview.png}
	\caption{\textbf{Full view of visualization platform.}
   The initial view of the visualization shows three panels: the laboratory-based DMS data, influenza variant frequencies in nature, and the 3D protein structure.
	}
	\label{overview}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=1.0\textwidth]{dms-brushed-tooltip.png}
	\caption{\textbf{Visualization of differential selection data generated from DMS experiments.}
  The laboratory data can be explored and zoomed in on by brushing the context panel below the primary panel, giving both detail and overview to the user.
Sites that are colored black in this panel represent sites for which there is natural frequency data.
For all grey sites, only one variant of influenza at that site is found to circulate naturally.
Mousing over a tip brings up details on demand, and clicking on a data point selects that site for visualization in the other panels, and colors the point according to its maximal differential selection value.
	}
	\label{dms-data}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=1.0\textwidth]{frequencies-panel.png}
	\caption{\textbf{Frequencies of influenza variants in nature.}
  Once a site has been selected from the laboratory data panel, this panel shows the frequencies of influenza viruses with different amino acids at that site over the last three years.
In this example, at site 193 we see that there are two variants circulating, one has a phenylalanine (F) at this site, while the other has a serine (S).
We can see that the phenyalanine variant is at higher frequencies, but that the serine variant is rising in frequency.
  }
	\label{frequencies-panel}
\end{figure}

@@

\begin{figure}[H]
	\includegraphics[width=1.0\textwidth]{3d-structure-panel.png}
	\caption{\textbf{3D protein structure with selected site highlighted.}
  3D protein structure panel, with the selected site colored according to its differential selection value.
The tooltip also provides information about which amino acid is at highest frequency at the site.
In this image, site 193 has a PHE, or phenylalanine.
  }
	\label{3d-structure}
\end{figure}

>>>>

%%
%% This is file `sample-sigchi.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigchi')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigchi.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigchi, nonacm]{acmart}

@@

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Unified visualization of seasonal influenza evolution in the laboratory and in nature.}

@@

Despite the value of these data, they are not intuitive to interpret.
Sequencing produces linear sequences, yet the protein conformation may have amino acid residues that are close together in 3D space that are quite distant from each other in linear sequence space.
Thus it can be hard to interpret how changes visible on a sequence map onto the protein structure.
Additionally, as DMS experiments represent controlled laboratory experiments looking at infection in cell culture, there are questions regarding how well DMS results reflect the patterns of influenza evolution that we observe in nature.
Here, we introduce an interactive visualization platform that allows unified display of DMS differential selection data, 3D protein structure, and information about amino acid preferences observed in nature, as unified by their common variable: site in the the sequence.

@@

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{viz-overview.png}
	\caption{\textbf{Full view of visualization platform.}
   The initial view of the visualization shows three panels: the laboratory-based DMS data, influenza variant frequencies in nature, and the 3D protein structure.
	}
	\label{overview}
\end{figure}

@@

This visualization platform links three data views together by their common shared element, site in the protein.
The three data streams are: 1) DMS differential selection value, which is a quantitative measure of the preference for a certain amino acid at a specific site in the protein, 2) the natural amino acid frequencies of a site, a quantitative variable the describes what proportion of all sampled strains of influenza at time T have a specific amino acid at a site, and 3) the 3D structure of protein itself, which takes linear sequence space and folds that up to reflect where certain sites of the sequence fall on the protein structure.
We describe design decisions about how to visualize these data streams below.
Our visualization was implemented using D3 \cite{bostock2011d3}

@@

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{dms-brushed-tooltip.png}
	\caption{\textbf{Visualization of differential selection data generated from DMS experiments.}
  The laboratory data can be explored and zoomed in on by brushing the context panel below the primary panel, giving both detail and overview to the user.
Sites that are colored black in this panel represent sites for which there is natural frequency data.
For all grey sites, only one variant of influenza at that site is found to circulate naturally.
Mousing over a tip brings up details on demand, and clicking on a data point selects that site for visualization in the other panels, and colors the point according to its maximal differential selection value.
	}
	\label{dms-data}
\end{figure}

@@

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{frequencies-panel.png}
	\caption{\textbf{Frequencies of influenza variants in nature.}
  Once a site has been selected from the laboratory data panel, this panel shows the frequencies of influenza viruses with different amino acids at that site over the last three years.
In this example, at site 193 we see that there are two variants circulating, one has a phenylalanine (F) at this site, while the other has a serine (S).
We can see that the phenyalanine variant is at higher frequencies, but that the serine variant is rising in frequency.
  }
	\label{frequencies-panel}
\end{figure}

@@

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{3d-structure-panel.png}
	\caption{\textbf{3D protein structure with selected site highlighted.}
  3D protein structure panel, with the selected site colored according to its differential selection value.
The tooltip also provides information about which amino acid is at highest frequency at the site.
In this image, site 193 has a PHE, or phenylalanine.
  }
	\label{3d-structure}
\end{figure}
